### 2024-08-30 draft.qmd
20240830 - draft and player eval

There is no one "right" draft strategy.
As noted by @Lee2022 in their analysis of fantasy drafts, the effectiveness of any draft strategy depends on the strategies of the other managers in the league.
Sometimes it works best to "zig" when everyone else is "zagging".
For instance, if you notice that everyone else is drafting Wide Receivers, this may mean that other managers are over-valuing Wide Receivers, and this could be a nice opportunity to draft a Running Back for good value.

@@

@sec-fantasyValueSources describes where to obtain [aggregated rankings](#sec-fantasyValueRankings), [aggregated projections](#sec-fantasyValueProjections), [ADP](#sec-fantasyValueADPAAV), and [AAV](#sec-fantasyValueADPAAV) data.

An important concept in the draft is "[dropoff](#sec-fantasyValueDropoff)", which is described in @sec-fantasyValueDropoff.
[*Dropoff*](#sec-fantasyValueDropoff) at a given position, is the difference—in terms of projected fantasy points—between (a) the best available player remaining at that position and (b) the second-best available player remaining at that position.
If there is a bigger [dropoff](#sec-fantasyValueDropoff) at a given position, there may be greater value in drafting the top player from that position.
For instance, consider the following scenario:
"Quarterback A" is projected to score 325 points, and "Quarterback B" is projected to score 320 points.
"Tight End A" is projected to score 230 points, and "Tight End B" is projected to score 150 points.
In this example, there is a much greater [dropoff](#sec-fantasyValueDropoff) for Tight Ends than there is for Quarterbacks.
Thus, even though "Quarterback A" is projected to score more points than "Tight End A", "Tight End A" may be more valuable because there is still a good Quarterback available if someone else drafts "Quarterback A".

Another important concept is a player's [value over a typical replacement player](#sec-fantasyValueVORP) at that position (shortened to "value over replacement player"; VORP), which is described in @sec-fantasyValueVORP.

@@

@@

>>>>

There is no one "right" draft strategy.
As noted by @Lee2022 in their analysis of fantasy drafts, the effectiveness of any draft strategy depends on the strategies of the other managers in the league.
Sometimes it works best to "zig" when everyone else is "zagging".
For instance, @Lee2022 found that the most successful draft strategies in terms of roster composition (i.e., the number of players for each position) were not the most common draft strategies.
For instance, if you notice that everyone else is drafting Wide Receivers, this may mean that other managers are over-valuing Wide Receivers, and this could be a nice opportunity to draft a Running Back for good value.

@@

As described in @sec-wisdomOfCrowd, there can be benefits of leveraging the [wisdom of the crowd](#sec-wisdomOfCrowd) by using rankings or projections that are averaged across many people and perspectives.
@sec-fantasyValueSources describes where to obtain [aggregated rankings](#sec-fantasyValueRankings), [aggregated projections](#sec-fantasyValueProjections), [ADP](#sec-fantasyValueADPAAV), and [AAV](#sec-fantasyValueADPAAV) data.
As described in @sec-judgmentVsActuarial, there are also benefits to using the [actuarial approach](#sec-judgmentVsActuarial) to prediction rather than (merely) using judgment.

It is not sufficient to compare players in terms of projected points because some positions have more depth than other positions.
Some positions show *positional scarcity*—that is, a limited number of high performing players.

An important concept in the draft is "[dropoff](#sec-fantasyValueDropoff)", which is described in @sec-fantasyValueDropoff.
[*Dropoff*](#sec-fantasyValueDropoff) at a given position, is the difference—in terms of projected fantasy points—between (a) the best available player remaining at that position and (b) the second-best available player remaining at that position.
If there is a bigger [dropoff](#sec-fantasyValueDropoff) at a given position, there may be greater value in drafting the top player from that position.
For instance, consider the following scenario:
"Quarterback A" is projected to score 325 points, and "Quarterback B" is projected to score 320 points.
"Tight End A" is projected to score 230 points, and "Tight End B" is projected to score 150 points.
In this example, there is a much greater [dropoff](#sec-fantasyValueDropoff) for Tight Ends than there is for Quarterbacks.
Thus, even though "Quarterback A" is projected to score more points than "Tight End A", "Tight End A" may be more valuable, relatively, because there is still a good Quarterback available if someone else drafts "Quarterback A".
In general, Kickers and Defenses tend to have the lowest dropoff (i.e., the lowest expected drop in fantasy points) by positional rank, so it makes sense to draft Kickers and Defenses late in the draft [@Lee2022].
Defenses, in particular, appear to be among the least predictable of the positions [@Lee2022].

Another important concept is a player's [value over a typical replacement player](#sec-fantasyValueVORP) at that position (shortened to "value over replacement player"; VORP), which is described in @sec-fantasyValueVORP.
A player's [value over a typical replacement player](#sec-fantasyValueVORP) provides a way to more fairly compare (and thus rank) players across different positions.

@@

In general, the most consequential decisions tend to be those made early in the draft regarding the top projected players [@Lee2022].
That is, the earlier selections tend to have the greatest impact on the success of one's fantasy season.
So, make sure to spend time in your draft preparation to identify the players you want to select early in the draft.
Moreover, some teams like to hedge their bet on their top Running Backs in the case that the player were to get injured, by drafting the player's backup, a strategy known as *handcuffing*.
However, there is not strong evidence that handcuffing leads to better outcomes [@Lee2022].

In general, @Lee2022 found that teams that drafted more Running Backs and Wide Receivers tended to outperform other teams.

@@

Herding behavior is common in snake drafts.
As described in @sec-wisdomOfCrowd, herding occurs when people align their behavior with others.
For instance, @Lee2022 found evidence that when Quarterbacks, Kickers, and Defenses were drafted by one team, subsequent teams were more likely to draft a player of that position.
The same was not the case for Running Backs, Wide Receivers, and Tight Ends.

### 2024-08-30 player-evaluation.qmd
20240829 - Vegas lines
20240830 - draft and player eval

@@

To reduce some of the bias due to a given source, some services aggregate projections across sources, consistent with a "wisdom of the crowd" approach.
Projections that are aggregated across sources are provided by [Fantasy Football Analytics](https://fantasyfootballanalytics.net) (see the webapp here: <https://apps.fantasyfootballanalytics.net>) and by [FantasyPros](https://www.fantasypros.com/nfl/auction-values/calculator.php).

>>>>

## Availability {#sec-evalAvailability}

In addition to [injuries](#sec-evalHealth), other factors can affect a player's availability, including league suspensions, for instance, due to violating the league's conduct policy.

@@

To reduce some of the bias due to a given source, some services aggregate projections across sources, consistent with a "wisdom of the crowd" approach.
Projections that are aggregated across sources are provided by [Fantasy Football Analytics](https://fantasyfootballanalytics.net) (see the webapp here: <https://apps.fantasyfootballanalytics.net>) and by [FantasyPros](https://www.fantasypros.com/nfl/auction-values/calculator.php).
Moreover, sites also provide season-long player futures for the over/under betting lines, which are based on Vegas lines and how people bet: <https://www.rotowire.com/betting/nfl/player-futures.php>.
